---
title: "PA1_template"
author: "Igor Vasyutin"
date: "08 03 2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Reproducible Research: Course Project 1**

### Loading and preprocessing the data

Loading data
```{r}
activity <- read.csv("activity.csv")
head(activity)
```

Convert date into date format
```{r}
activity$date <- as.Date(activity$date)
```

### What is mean total number of steps taken per day?

Calculate mean per day
```{r}
meanday <- aggregate(steps ~ date, data = activity, mean)
head(meanday)
```

print hist, mean and median for mean steps per day
```{r}
hist(meanday$steps,breaks = 20)
mean(meanday$steps,na.rm=TRUE)
median(meanday$steps,na.rm=TRUE)
```

### What is the average daily activity pattern?

Calculate mean per interval and make a plot of average activity
```{r}
meanperiod <- aggregate(steps ~ interval, data=activity, mean)
head(meanperiod)
plot(steps ~ interval, type = "l", data=meanperiod)
```

5-minutes interval with max number of steps
```{r}
meanperiod[meanperiod$steps==max(meanperiod$steps, na.rm=TRUE),1]
```

### Imputing missing values

How many NA's:
```{r}
sum(is.na(activity$steps))
sum(is.na(activity$steps))/length(activity$steps)
```

Filling NA's with average for an interval across all days
```{r}
activity.imputed <- activity
for (i in 1:nrow(activity.imputed)){
    if (is.na(activity.imputed$steps[i])){
        intrvl <- activity.imputed$interval[i]
        activity.imputed$steps[i]<-meanperiod[meanperiod$interval==intrvl,2]
    }
}
sum(is.na(activity.imputed$steps))
```

print hist, mean and median for mean steps per day on imputed dataset
```{r}
meanday.imputed <- aggregate(steps ~ date, data = activity.imputed, mean)
hist(meanday.imputed$steps,breaks = 20)
mean(meanday.imputed$steps)
median(meanday.imputed$steps)
```

### Are there differences in activity patterns between weekdays and weekends?

Make a vector of weekdays and weekends
```{r}
activity.imputed$weekdays<-weekdays(activity.imputed$date,abbreviate = TRUE)
activity.imputed$weekdays[activity.imputed$weekdays == "Сб" | activity.imputed$weekdays == "Вс"] <- "WE"
activity.imputed$weekdays[activity.imputed$weekdays != "WE"] <- "WD"
```

Make a plot comparing average activity at weekdays and weekends
```{r}
meanperiod.imputed.wd <- aggregate(steps ~ interval, data = activity.imputed[activity.imputed$weekdays=="WD",], mean)
meanperiod.imputed.we <- aggregate(steps ~ interval, data = activity.imputed[activity.imputed$weekdays=="WE",], mean)
with(meanperiod.imputed.wd, plot(interval, steps, type="l"))
with(meanperiod.imputed.we, lines(interval, steps, col = "red"))
```